---
title: "PA1_template"
author: "Brayam Pineda"
output: html_document
---

## Objective

This file contain the solve to the first project of the Reproducible Research 
course in Coursera offered by Johns Hopkins University.For it we will to analyze
the activity dataset that you can find in [Activity monitoring data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)

In short, we need to complete five general assignments, they are:

- Loading and preprocessing the data
- What is mean total number of steps taken per day?
- What is the average daily activity pattern?
- Imputing missing values
- Are there differences in activity patterns between weekdays and weekends?

So let's go

## 1. Loading and preprocessing the data

```{r, echo=TRUE}
data <- read.csv("activity.csv")
head(data)
```

That is not too informative, so we are going to use the *str()* function

```{r}
str(data)
```
We can see that the class of the columns is not the best, so we are going to
solve this

```{r}
data$steps <- as.numeric(data$steps)
data$date <- as.Date(data$date)
data$interval <- as.numeric(data$interval)
str(data)
```
Is a good practice explore a little more the data before start with the analyze,
a good way is through the *summary()* function


```{r}
summary(data)
```

## 2. What is mean total number of steps taken per day?

First we need to transform the data base

```{r, echo=TRUE, message=FALSE}
library(dplyr)
data_v2 <- data %>% group_by(date) %>% summarise(Day = sum(steps, na.rm = FALSE))
data_v2 <- data_v2[!is.na(data_v2$Day), ]
head(data_v2)
```

After that, we can do a histogram to see how the data is behavior

```{r, , message=FALSE}
library(ggplot2)
ggplot(data = data_v2, aes(x = Day)) + geom_histogram(color="#e9ecef", fill="#69b3a2") +
  theme_bw() + xlab("Count of the steps by Day") + ylab("")
```
By last, the task request us for calculate the *median* and *mean* of the total of
steps, the easiest way is by the *summary* function and we can to report that
result in a table

```{r}
summary(data_v2$Day)
```



