---
title: "PA1_template"
author: "Brayam Pineda"
output:
  html_document: default
  pdf_document: default
---

## Objective

This file contain the solve to the first project of the Reproducible Research 
course in Coursera offered by Johns Hopkins University.For it we will to analyze
the activity dataset that you can find in [Activity monitoring data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)

In short, we need to complete five general assignments, they are:

- Loading and preprocessing the data
- What is mean total number of steps taken per day?
- What is the average daily activity pattern?
- Imputing missing values
- Are there differences in activity patterns between weekdays and weekends?

So let's go

## 1. Loading and preprocessing the data

```{r, echo=TRUE}
data <- read.csv("activity.csv")
head(data)
```

That is not too informative, so we are going to use the *str()* function

```{r}
str(data)
```
We can see that the class of the columns is not the best, so we are going to
solve this

```{r}
data$steps <- as.numeric(data$steps)
data$date <- as.Date(data$date)
data$interval <- as.numeric(data$interval)
str(data)
```
Is a good practice explore a little more the data before start with the analyze,
a good way is through the *summary()* function


```{r}
summary(data)
```

## 2. What is mean total number of steps taken per day?

First we need to transform the data base

```{r, echo=TRUE, message=FALSE}
library(dplyr)
data_v2 <- data %>% group_by(date) %>% summarise(Day = sum(steps, na.rm = FALSE))
data_v2 <- data_v2[!is.na(data_v2$Day), ]
head(data_v2)
```

After that, we can do a histogram to see how the data is behavior

```{r, , message=FALSE}
library(ggplot2)
ggplot(data = data_v2, aes(x = Day)) + geom_histogram(color="#e9ecef", fill="#69b3a2") +
  theme_bw() + xlab("Count of the steps by Day") + ylab("")
```

By last, the task request us for calculate the *median* and *mean* of the total of
steps, the easiest way is by the *summary* function and we can to report that
result in a table

```{r}
summary(data_v2$Day)
```
## 3. What is the average daily activity pattern?

- Make a time series plot (i.e. \color{red}{\verb|type = "l"|}type = "l") of the
  5-minute interval (x-axis) and the average number of steps taken, averaged 
  across all days (y-axis)
  
```{r}
data_v3 <- data %>% group_by(interval) %>% summarise("Average of steps" = mean(steps, na.rm = TRUE))
ggplot(data = data_v3, aes(x = interval, y = `Average of steps`)) + 
  geom_line() + theme_bw() +
  xlab("Intervals of 5 minutes")
```

- Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
max <- data_v3 %>% filter(`Average of steps` == max(`Average of steps`))
knitr::kable(max, caption = ""
       , align = c('c')
       , col.names = c("Interval", "Average of steps")
       , row.names = FALSE
       , justify = 'lcrrcl')
```

In graphic it will be

```{r}
ggplot(data = data_v3, aes(x = interval, y = `Average of steps`)) + 
  geom_line() + theme_bw() + geom_vline(aes(xintercept = 835, colour = "red"), size = 1, show.legend = FALSE) +
  xlab("Intervals of 5 minutes")
```

## 4. Imputing missing values

- Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with \color{red}{\verb|NA|}NAs)

```{r}
sum(is.na(data$steps))
```

- Devise a strategy for filling in all of the missing values in the dataset.

```{r message=FALSE, results='hide', , warning=FALSE}
library(mice)
data_v4 <- data
imputed_data1 <- mice(data_v4)
complete_data1 <- mice::complete(imputed_data1)
```

- Make a histogram of the total number of steps taken each day and Calculate and
  report the mean and median total number of steps taken per day. Do these values
  differ from the estimates from the first part of the assignment? What is the
  impact of imputing missing data on the estimates of the total daily number of steps?

```{r, message=FALSE}
complete_data2 <- complete_data1 %>% group_by(date) %>% summarise(Steps = sum(steps, na.rm = TRUE))
ggplot(data = complete_data2, aes(x = Steps)) + geom_histogram(color="#e9ecef", fill="#69b3a2") +
  theme_bw() + xlab("Count of the steps by Day") + ylab("") 

```

Recall the other graphic

```{r, echo=FALSE, message=FALSE}
ggplot(data = data_v2, aes(x = Day)) + geom_histogram(color="#e9ecef", fill="#69b3a2") +
  theme_bw() + xlab("Count of the steps by Day") + ylab("")
```

Although based on the graphic it appear the data have not had important changes 
the task require us calculate the median and mean

```{r}
summary(data_v2$Day)
summary(complete_data2$Steps)
```

There are notable differences between both


